---
title: "Predicting Exercise Notes"
author: "Helen"
date: "7 February 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warnings = FALSE, cache = TRUE)
```

## Plan

```{r data and libraries}
library(caret)
library(gbm)
library(e1071)
training <- read.csv("~/Personal/Coursera/Practical Machine Learning/pml-training.csv")
```

Use k-fold validation to test different models and tune parameters.

Split the sample into folds using stratification to ensure same numbers of each type of error. Default is 10 and apparently there is evidence that 10 is good.

```{r folds}
set.seed(134)
inTrain <- createFolds(training$classe) ## Default is 10
```

## Exploratory plots

```{r expplots}
expplots <- function(data, first, last, name){
  PDFpath <- paste("~/Personal/Coursera/Practical Machine Learning/PracticalML/", name, ".pdf")
  pdf(file = PDFpath)
  for (i in first:last){
    g <- ggplot(data, aes(x = classe, y = data[, i])) + geom_violin() + ylab(names(data[i])) + ggtitle(paste("Column "), i)
    print(g)
  }
  dev.off()
}

expplots(8, 160, "expplots")
```

This produced a pdf of lots of plots. The distributions look quite different between the classes for columns:
32, 50, 51, 52, 56, 60, 61, 63, 65, 66, 67, 75, 77, 82, 83, 84, 104, 105, 107, 117, 131, 154

And a bit different for columns:
40, 46, 49, 54, 55, 64, 76, 78, 80, 96, 100, 102, 108,  110, 121, 122, 123, 134, 137, 155

```{r 1st useful data}
training.1 <- training[, c(1, 32, 50, 51, 52, 56, 60, 61,63, 65, 66, 67, 75, 77, 82, 83, 84, 104, 105, 107, 117, 131, 154, 160)]
```

Look at these in combination with each other.

```{r 1st data scatter}
expplots2 <- function(data, first, last, name){
  PDFpath <- paste("~/Personal/Coursera/Practical Machine Learning/PracticalML/", name, ".pdf", sep = "")
  pdf(file = PDFpath)
  a <- ceiling(sqrt(last-first+1))
  for (i in first:last){
    plots <- list()
    for (j in first:last){
      thing <- ggplotGrob(ggplot(data, aes(data[, i], data[, j], colour = data$classe)) + geom_point() + xlab(names(data[, i])) + ylab(names(data[, j])) + ggtitle(paste("Columns ", i, " and ", j, sep = "")) +guides(colour = "none"))
      plots[[j - first + 1]] <- thing
  }
  do.call(grid.arrange, plots)
  }
  dev.off()
}
```